Leader functions and qualities across 59 cultures: an exploratory analysis

### 2019-12-03 leadership_across_cultures_contexts.Rmd
Additional minor edits.

>Their domestic policy may be described as a communism modified by the authority, more or less nominal, of the chief. The head chief of a tribe is called maia igla and the elders, or sub-chiefs, i.e. those in authority over each community, consisting of from 20 to 50 individuals, maicla . The head chief, who usually resides at a permanent encampment, has authority over all the sub-chiefs, but his power, like theirs, is very limited. It is exercised mainly in organising meetings between the various communities belonging to his tribe, and in exerting influence in all questions affecting the welfare of his followers. It is the chief alone, as may be supposed, who directs the movements of a party while on hunting and fishing expeditions, or when migrating. It is usually through his intervention that disputes are settled, but he possesses no power to punish or enforce obedience to his wishes, it being left to all alike to take the law into their own hands when aggrieved. The aryoto and eremtaga in each tribe have their own head chief, who are independent the one of the other."

@@

>"Bravery may be considered as a means of acquiring status in what Freeman has identified as an egalitarian society. The leader of a longhouse, the tuai rumah, is elected by both men and women of the longhouse for his skills in oratory, warfare, and farming, and his familiarity with adat , or customary law, in order to resolve disputes (Freeman 1970:111). Although usually a man, women tuai rumah are not unknown. The tuai rumah must inspire the confidence of his followers and exemplify the qualities of bravery, endurance, and determination that are essential for survival. Skill in farming and warfare is important in the pioneering context, in which the tuai rumah leads a group to establish a new settlement. The tuai rumah has no authority to command others, but is empowered to resolve disputes through influence and by reference to custom."

@@

>Wanano chieftainship entails control over three major aspects of village life: fishing resources, labor, and ritual. A mahsa wami must hold sway over all three areas to lead effectively. Inability to harvest and redistribute fish, for example, will make it difficult for the mahsa wami to mobilize and coordinate labor, Control, then, implies the obligation to redistribute. The chief mobilizes and coordinates communal labor, carefully balancing collective projects against individual work such as gardening, so that collective and individual labor do not conflict. I observed several kinds of community projects in Wanano villages, including the construction and maintenance of houses, the clearing and burning of gardens, and the intensive preparations for intersib po?oa exchange ceremonies.

@@

One interesting pattern that has not been widely discussed in the evolutionary literature on leadership is the role of shamans and their intersection, in our results, with the dominant style of leadership. Our social costs cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* abilities variable. @winkelman1987cross found that in 47% of societies sampled from the HRAF ($n=43$) shamans were associated with providing decision-making services and associated with military power. Providing ritual functions and military leadership are common (documented in over 50% of societies here) functions leaders serve (see Figure \@ref(fig:varsupport)). Widespread preferences for egalitarianism and social equality are likely strong drivers of leader preferences across diverse societies [@cheng_dominance_2019], but shamanism and ritual based contexts represent a social domain in which ideologies support dominance-based leadership.

We also found a link between prestige-based leadership, which is associated with deference towards individuals with special skills, high intelligence, and accomplishments [@cheng_dominance_2019; @van_vugt_dual_2019; @henrich_evolution_2001], and charismatic leadership, which is associated with oratory skills, high charisma, confidence, and interpersonal skills [@grabo_charismatic_2016; @howell_role_2005]. See, Figure \@ref(fig:cluster)A, *Ability to provide social benefits* cluster. This suggests that these leadership strategies might be closely related across a broad range of societies and social contexts, or part of a broader strategy of social influence based on the provisioning of various benefits.

>>>>

>Their domestic policy may be described as a communism modified by the authority, more or less nominal, of the chief. The head chief of a tribe is called maia igla and the elders, or sub-chiefs, i.e. those in authority over each community, consisting of from 20 to 50 individuals, maicla . The head chief, who usually resides at a permanent encampment, has authority over all the sub-chiefs, but his power, like theirs, is very limited. It is exercised mainly in organising meetings between the various communities belonging to his tribe, and in exerting influence in all questions affecting the welfare of his followers. It is the chief alone, as may be supposed, who directs the movements of a party while on hunting and fishing expeditions, or when migrating. It is usually through his intervention that disputes are settled, but he possesses no power to punish or enforce obedience to his wishes, it being left to all alike to take the law into their own hands when aggrieved. The aryoto and eremtaga in each tribe have their own head chief, who are independent the one of the other.

@@

>Bravery may be considered as a means of acquiring status in what Freeman has identified as an egalitarian society. The leader of a longhouse, the tuai rumah, is elected by both men and women of the longhouse for his skills in oratory, warfare, and farming, and his familiarity with adat , or customary law, in order to resolve disputes (Freeman 1970:111). Although usually a man, women tuai rumah are not unknown. The tuai rumah must inspire the confidence of his followers and exemplify the qualities of bravery, endurance, and determination that are essential for survival. Skill in farming and warfare is important in the pioneering context, in which the tuai rumah leads a group to establish a new settlement. The tuai rumah has no authority to command others, but is empowered to resolve disputes through influence and by reference to custom.

@@

>Wanano chieftainship entails control over three major aspects of village life: fishing resources, labor, and ritual. A mahsa wami must hold sway over all three areas to lead effectively. Inability to harvest and redistribute fish, for example, will make it difficult for the mahsa wami to mobilize and coordinate labor. Control, then, implies the obligation to redistribute. The chief mobilizes and coordinates communal labor, carefully balancing collective projects against individual work such as gardening, so that collective and individual labor do not conflict. I observed several kinds of community projects in Wanano villages, including the construction and maintenance of houses, the clearing and burning of gardens, and the intensive preparations for intersib po?oa exchange ceremonies.

@@

One interesting pattern that has not been widely discussed in the evolutionary literature on leadership is the role of shamans and their intersection, in our results, with the dominant style of leadership. Our social costs cluster (Figure \@ref(fig:cluster)A) includes the *Supernatural* abilities variable. @winkelman1987cross found that in 47% of societies sampled from the HRAF ($n = 43$) shamans were associated with providing decision-making services and associated with military power. Providing ritual functions and military leadership are common (documented in over 50% of societies here) functions leaders serve (see Figure \@ref(fig:varsupport)). Widespread preferences for egalitarianism and social equality are likely strong drivers of leader preferences across diverse societies [@cheng_dominance_2019], but shamanism and ritual based contexts represent a social domain in which ideologies support dominance-based leadership.

We also found a link between prestige-based leadership, which is associated with deference towards accomplished individuals with expertise, special skills, and high intelligence [@cheng_dominance_2019; @henrich_evolution_2001], and charismatic leadership, which is associated with oratory skills, charismatic personality, confidence, and interpersonal skills [@grabo_charismatic_2016; @howell_role_2005]. See, Figure \@ref(fig:cluster)A, *Ability to provide social benefits* cluster. This suggests that these leadership strategies might be closely related across a broad range of societies and social contexts, or part of a broader strategy of social influence based on the provisioning of various benefits.
